Embedding Intuitionistic into Classical Logic

### 2022-07-14 CSL/lipics-v2021-EIICL.tex
all comments for the technical sections added

It is well known that a polynomial time translation between QBF and IPC must exists, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}. However giving an explicit translation from IPC to QBF can still be useful as good solvers exists for QBF but for IPC not so much. The translation from the last section will serve as a useful tool for this. We will not explain the semantics of QBF in depth. Intuitively just think that the interpretation of existentially quantified variables may depend on variables quantified above them.

Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P$$
and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

@@

\begin{lemma}
	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
\end{lemma}
\begin{proof}
	Suppose $\varphi$ is not intuitionistically valid, i.e. there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\epsilon$. For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$. Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.

@@

Since we have not discussed QBF in detail the above argument might be a bit difficult to follow, in particular the second part, but is simple in essence. We shall demonstrate with help of a example.

\begin{example}
	Consider the formula $\varphi = (((A\to B)\to C)\wedge((B\to A)\to C))\to C$, which is a non-intuitionistic tautology. Its associated QBF is $$\varphi^Q = \exists A^0B^0C^0\forall F_{\psi_1}^1F_{\psi_2}^1 \exists A^1B^1C^1\forall F_{\psi_1}^2F_{\psi_2}^2 \exists A^2B^2C^2\psi'$$where $\psi_1 := (A\to B)\to C, \psi_2 = (B\to A)\to C$ and $$\psi' = \bigwedge\left\{\begin{matrix}
		F^1_{\psi_1}\vee F^1_{\psi_2}, \neg(F^1_{\psi_1}\wedge F^1_{\psi_2}), F^2_{\psi_1}\vee F^2_{\psi_2},\neg(F^2_{\psi_1}\wedge F^2_{\psi_2}),\\F^1_{\psi_1}\to\neg F^2_{\psi_1},F^1_{\psi_2}\to\neg F^2_{\psi_2}, A^0\to A^1,A^1\to A^2,
		\\B^0\to B^1, B^1\to B^2, C^0\to C^1, C^1\to C^2,\neg C^0, \\
		F^1_{\psi_1}\to (A^1\to B^1)\to C^0, F^1_{\psi_2}\to(B^1\to A^1)\to C^0\\
		F^2_{\psi_1}\to (A^2\to B^2)\to C^1, F^2_{\psi_2}\to(B^2\to A^2)\to C^1
	\end{matrix}\right\}$$
	This QBF is satisfiable: Have $P^0 = 0$ for all atoms $P$, always $P^2 = 1$ for all atoms $P$ and $C^1 = 1$. Furthermore have $A^1 = 1$ iff $F^1_{\psi_1} = 1$ and $B^1 = 1$ iff $F^1_{\psi_2} = 1$. The construction of an counter-model to $\varphi^\#$ then proceeds as follows:
	\begin{itemize}
		\item First have $P^\epsilon = 0$ for each propositional variable $P$.
		\item For $P^{\psi_1}$ we consider what happens when we choose $F^1_{\psi_1}$ to be true, i.e. we have $A^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
		\item For $P^{\psi_2}$ we consider what happens when we choose $F^1_{\psi_2}$ to be true, i.e. we have $B^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
		\item Finally for $P^{\psi_1\psi_2} = P^{\psi_2\psi_1} = 1$ for all propositional variables $P$ as $P^2$ is true regardless of choice for the $F^i_\psi$.
	\end{itemize}
\end{example}

Note that the last set of existentially quantified variables is superfluous because there is only one assignment which even has the chance to falsify $\varphi^Q$, namely that which assigns $1$ to the only $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$. That is in our final formula we can replace $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^i$. With that we get the following:

\begin{corollary}
	Let $N$-INTINVALID be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid. Then $N$-INTINVALID is in $\Sigma_{2N-1}^P$. The dual problem $N$-INTVALID is in $\Pi_{2N-1}^P$.
\end{corollary}

@@

	The other direction is a bit more involved. Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$. In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$. Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$. Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$ and extend $\sim$ by having $[w] = \{w\}$ for $w\in W_0$. Now have
	\begin{itemize}
		\item $s^I = w_0$.
		\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
		\item $f^I(m_1\dots m_n) =\begin{cases}
			 f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
			 w_0, & \text{otherwise.}
		\end{cases}$
		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
			A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
			\top, & \text{otherwise.}
		\end{cases}$
	\end{itemize}
	One easily verifies that these are well-defined. Let us now briefly check that this indeed defines a counter-model. First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is. Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified. All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$. Consider e.g. the case where $\psi\in\mathcal S$ is of the form
	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
	We have to show that
	$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
	$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false. For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$. Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$. Note that the above is false in particular for $w\in M_0$ with $u\preceq w$. But this would imply
	$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
\end{proof}

@@

As in the predicate case we now show the existence of a canonical counter-model, if a counter-model exists. Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$. As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$ with
$$\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$$ and
$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$$ with $$\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$$
and in each other formula in $\mathcal S^\circ$ removing the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
As in the previous section we can define a tree-counter model for this.
\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. For $\psi = \forall\vec t\psi'\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff $(A_\psi^I(\vec t_A, u)\to B_\psi^I(\vec t_B, u))[\vec a/\vec t]$ is false or $C_\psi^I(\vec t_C, u)[\vec a/\vec t]$ is true.
	For $\psi = \forall\vec t\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff there is $a\in M_u$ such that $A^{I[a/x]}_\psi(\vec t_A)[\vec a/\vec t]$ is false or $B_\psi^I(\vec t_B)[\vec a/\vec t]$ is true. For $\psi\in\mathcal F$ define $$g_{\vec a, \psi} : M\to M, u\mapsto\begin{cases}
		u, &\text{ if $\psi$ is fulfilled in $u$ with $\vec a$,}\\
		f^I_\psi(\vec a, u), &\text{ else.}		
	\end{cases}$$		
	Define a $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ are the sequences on $\{ \vec t, \psi\:|\:\psi\in \mathcal F, \vec t\in M^n\}$ without repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\to$ and without immediate repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\forall$.
		\item Interpret $\preceq$ as the prefix-order.
		\item Have $$f_\psi^{I_T}(x_1\dots x_n) = \begin{cases}
			x_1\dots x_n,&\text{if $x=x_n$,}\\
			x_1\dots x_nx,&\text{else.}			
		\end{cases}$$
		\item For propositional variables $P$ have $$P^{I_T}(x_1\dots x_n) = P^I(g_{x_n}(\dots(g_{x_1}(s^I))\dots))$$
	\end{itemize}
\end{definition}

\begin{lemma}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_\psi^I(\vec a, u)$ with $\vec a$ for all $u\in M, \psi\in\mathcal F$, $\vec a\in M^n$.
		\item If $\psi\in\mathcal F_\to$ is fulfilled at $u\in M$ with $\vec a$ then $\psi$ is fulfilled at all $w\geq u$ with $\vec a$.
	\end{enumerate}	
\end{lemma}

@@

Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$$ is sufficient. In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PO-encoding from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$. Define

$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#$$
then we have
\begin{theorem}
	$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
\end{theorem}
The complete translation is summarised in Theorem~\ref{fullFOtranslation}.

>>>>

It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}. 
However, giving an explicit translation from IPC to QBF still promises to be useful, allowing to leverage the progress in QBF solving.
The translation from the last section will serve as starting point for developing the translation in this section.
For lack of space, we do not define the semantics of QBF in this paper and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.

Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

@@

\begin{lemma}
	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
\end{lemma}
\begin{proof}
\todo{The proof contains some old notation at several places.}
\todo{Also the proof could be beautified a bit.}
Suppose $\varphi$ is not intuitionistically valid, i.e. there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\epsilon$.
For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$. Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.

@@

\begin{example}
\todo{I think this example is a candidate for the appendix.}
	Consider the formula $\varphi = (((A\to B)\to C)\wedge((B\to A)\to C))\to C$, which is a non-intuitionistic tautology. Its associated QBF is $$\varphi^Q = \exists A^0B^0C^0\forall F_{\psi_1}^1F_{\psi_2}^1 \exists A^1B^1C^1\forall F_{\psi_1}^2F_{\psi_2}^2 \exists A^2B^2C^2\psi'$$where $\psi_1 := (A\to B)\to C, \psi_2 = (B\to A)\to C$ and $$\psi' = \bigwedge\left\{\begin{matrix}
		F^1_{\psi_1}\vee F^1_{\psi_2}, \neg(F^1_{\psi_1}\wedge F^1_{\psi_2}), F^2_{\psi_1}\vee F^2_{\psi_2},\neg(F^2_{\psi_1}\wedge F^2_{\psi_2}),\\F^1_{\psi_1}\to\neg F^2_{\psi_1},F^1_{\psi_2}\to\neg F^2_{\psi_2}, A^0\to A^1,A^1\to A^2,
		\\B^0\to B^1, B^1\to B^2, C^0\to C^1, C^1\to C^2,\neg C^0, \\
		F^1_{\psi_1}\to (A^1\to B^1)\to C^0, F^1_{\psi_2}\to(B^1\to A^1)\to C^0\\
		F^2_{\psi_1}\to (A^2\to B^2)\to C^1, F^2_{\psi_2}\to(B^2\to A^2)\to C^1
	\end{matrix}\right\}$$
	This QBF is satisfiable: Have $P^0 = 0$ for all atoms $P$, always $P^2 = 1$ for all atoms $P$ and $C^1 = 1$. Furthermore have $A^1 = 1$ iff $F^1_{\psi_1} = 1$ and $B^1 = 1$ iff $F^1_{\psi_2} = 1$. The construction of an counter-model to $\varphi^\#$ then proceeds as follows:
	\begin{itemize}
		\item First have $P^\epsilon = 0$ for each propositional variable $P$.
		\item For $P^{\psi_1}$ we consider what happens when we choose $F^1_{\psi_1}$ to be true, i.e. we have $A^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
		\item For $P^{\psi_2}$ we consider what happens when we choose $F^1_{\psi_2}$ to be true, i.e. we have $B^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
		\item Finally for $P^{\psi_1\psi_2} = P^{\psi_2\psi_1} = 1$ for all propositional variables $P$ as $P^2$ is true regardless of choice for the $F^i_\psi$.
	\end{itemize}
\end{example}

Note that the last set of existentially\todo{universally?} quantified variables can be avoided because there is only one assignment that has the chance to falsify $\varphi^Q$, namely the assignment that assigns $1$ to the single $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$. Hence, in our final formula we can replace every $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^n$. With that we get the following:

\begin{corollary}
	Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid. Then $N$-\textbf{IntInvalid} is in $\Sigma_{2N-1}^P$. The dual problem $N$-\textbf{IntValid} is in $\Pi_{2N-1}^P$.
\end{corollary}

\todo{The size analysis of the QBF reduction should be stated in some theorem in this section}

@@

	The other direction is a bit more involved. Suppose we have a Kripke counter-model to $\varphi$ with Frame \todo{frame? Also make sure frame is defined} $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$. In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$. Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$. Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$ and extend $\sim$ by having $[w] = \{w\}$ for $w\in W_0$ \todo{why is the later needed? I don't think it is used}. 
Now have
	\begin{itemize}
		\item $s^I = w_0$.
		\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
		\item $f^I(m_1\dots m_n) =\begin{cases}
			 f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
			 w_0, & \text{otherwise.}
		\end{cases}$
		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
			A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
			\top, & \text{otherwise.}
		\end{cases}$
	\end{itemize}
	One easily verifies that these are well-defined. Let us now briefly check that this indeed defines a counter-model. First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is. Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified. All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$. Consider e.g. the case where $\psi\in\mathcal S$ is of the form
	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
	We have to show that
	$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
	$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false. For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$. Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$. Note that the above is false in particular for $w\in M_0$ with $u\preceq w$. But this would imply
	$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
\end{proof}

@@

As in the predicate case we now show the existence of a canonical counter-model, if a counter-model exists. Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$. As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$ with
$$\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$$ and
$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$$ with $$\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$$
and in each other formula in $\mathcal S^\circ$ removing the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
As in the previous section we can define a tree-counter model for this.
\todo[inline]{Important: Why can't we have Theorem~\ref{thm:fo-reduction} directly, without Def.~\ref{def:fo-fulfilled} and Lemma~\ref{thm:fo-countermodel-reduction}? They don't seem to be needed.
However, Lemma~\ref{thm:fo-countermodel-reduction} could still be used as in the propositional case to eliminate the function symbols for the implications at the cost of introducing additional predicates. 
One could also add the idempotence as axioms in order to aid the solver. 
We could mentioned that as a justification for Lemma~\ref{thm:fo-countermodel-reduction}.
}

\begin{definition}
\label{def:fo-fulfilled}
\todo{this definition still uses variables $t$, please use a different letter}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. For $\psi = \forall\vec t\psi'\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff $(A_\psi^I(\vec t_A, u)\to B_\psi^I(\vec t_B, u))[\vec a/\vec t]$ is false or $C_\psi^I(\vec t_C, u)[\vec a/\vec t]$ is true.
	For $\psi = \forall\vec t\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff there is $a\in M_u$ such that $A^{I[a/x]}_\psi(\vec t_A)[\vec a/\vec t]$ is false or $B_\psi^I(\vec t_B)[\vec a/\vec t]$ is true. For $\psi\in\mathcal F$ define $$g_{\vec a, \psi} : M\to M, u\mapsto\begin{cases}
		u, &\text{ if $\psi$ is fulfilled in $u$ with $\vec a$,}\\
		f^I_\psi(\vec a, u), &\text{ else.}		
	\end{cases}$$		
	Define a $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ are the sequences on $\{ \vec t, \psi\:|\:\psi\in \mathcal F, \vec t\in M^n\}$ without repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\to$ and without immediate repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\forall$. \todo{do you mean sequences of pairs here? Then it is probably better to use brackets as in $(\vec t, \psi)$}
		\item Interpret $\preceq$ as the prefix-order.
		\item Have $$f_\psi^{I_T}(x_1\dots x_n) = \begin{cases}
			x_1\dots x_n,&\text{if $x=x_n$,}\\
			x_1\dots x_nx,&\text{else.}			
		\end{cases}$$
		\item For propositional variables $P$ have $$P^{I_T}(x_1\dots x_n) = P^I(g_{x_n}(\dots(g_{x_1}(s^I))\dots))$$
	\end{itemize}
\end{definition}

\begin{lemma}
\label{thm:fo-countermodel-reduction}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_\psi^I(\vec a, u)$ with $\vec a$ for all $u\in M, \psi\in\mathcal F$, $\vec a\in M^n$.
		\item If $\psi\in\mathcal F_\to$ is fulfilled at $u\in M$ with $\vec a$ then $\psi$ is fulfilled at all $w\geq u$ with $\vec a$.
	\end{enumerate}	
\end{lemma}

@@

Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$$ is sufficient. In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PO-encoding \todo{PO-encoding used for the first time, say PartialOrder?}
from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$. Define

$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#$$
then we have
\begin{theorem}
\label{thm:fo-reduction}
	$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
\end{theorem}
The complete translation is summarised in Theorem~\ref{fullFOtranslation}.
